---
title: "Homework Week 7"
author: "Peter Gatica"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# library(gtools)
library(matlib)
```

1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly       distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the             distribution of Y .

$$ For\  1\leq j\leq k,m(j)=\frac{(k-j+1)^{n}-(k-j)^{n}}{k^{n}} $$

2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine
   has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every    ten years. (Include the probability statements and R Code for each part.).

a. What is the probability that the machine will fail after 8 years?. Provide also the
   expected value and standard deviation. Model as a geometric. (Hint: the probability is
   equivalent to not failing during the first 8 years..)
   
   $$ Probability\ Formula:  P(X=x)=(1-p)^{n-1}\ast p $$

```{r}
p=1/10
x=8
n=9
(prob_8y = (1-p)^(n-1) * p)
(dgeom(x,p))
```

$$ Expected\  Value\  Formula:\  E(X)=\frac{1}{p}\ or\ p^{-1} $$
```{r}
p=1/10
(EV=1/p)
(EV=p^-1)
```
$$
Standard\ Deviation\ Formula: sd=\sqrt{\frac{1-p}{p^{2}} } 
$$
```{r}
(sd = sqrt((1-p)/p^2))
```
b. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential.

$$ Probability\  as\  Exponential:P(X>=8)=e^{\frac{-k}{\mu } } $$
```{r}
k=8
lambda=1/10
mu=1/lambda
(prob_8y=exp((-k/(mu))))
```
$$ Expected\  Value:E[X]=u=\frac{1}{\lambda } =10,where\ lambda =\frac{1}{10} $$
```{r}
(EV=mu)
```

$$ Standard\  Deviation:sd=\sqrt{\frac{1}{{}^{\lambda^{2} }} }  $$
```{r}
(sd=sqrt(1/lambda^2))
```
c. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years)

$$ Probability:P(X>8)=1-p^{x}(1-p)^{n-x} $$
```{r}
(n=10)
(x=8)
(choose(n,x)*p^x*((1-p)^(n-x)))
```
$$ Expected\  Value:E[X]=np $$
```{r}
(n*p)
```
$$
Standard\ Deviation\ Formula: \sigma =\sqrt{npq} \  where\  q=1-p
$$
```{r}
q=1-p
(sd = sqrt(n*p*q))
```

d. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson.
$$ P(X=8)=\frac{{}\lambda^{x} e^{-\lambda }}{x!} \  Where\  \lambda =\frac{np}{t} $$
```{r}
t=1
x=8
(lambda=(n*p)/t)
(p_pois=((lambda^x)*(exp(-lambda)))/factorial(x))
```
$$ E[X]=\lambda $$
```{r}
(lambda)
```
$$ sd=\sqrt{\lambda } $$
```{r}
(sd=sqrt(lambda))
```



