---
title: "Assignment Week 10"
author: "Peter Gatica"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(expm)
library(markovchain)
library(diagram)
library(pracma)
```

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.
Find the probability that he wins 8 dollars before losing all of his money if

(a) he bets 1 dollar each time (timid strategy)
```{r}
p <- c(.6,0,.4)
bb <- c(1,rep(0,8),p,rep(0,7),p,rep(0,7),p,rep(0,7),p,rep(0,7),p,rep(0,7),p,rep(0,7),p,rep(0,8),1)
P_bb <- matrix(bb,9,9,byrow=TRUE)
```
##### Build the Markov Chain matrix
```{r}
(Pbb <-new('markovchain', transitionMatrix = P_bb, states = c('0','1','2','3','4','5','6','7','8'),name = 'Bet for Bail'))
```
##### Identify transient states
```{r}
# Determine transient states
(transientStates(Pbb))
```
##### Identify absorbing states
```{r}
# determine absorbing states
(absorbingStates(Pbb))
```
##### Create the Markov Chain in Conanical form
```{r}
p <- c(.6,0,.4)
bb <- c(0,.4,rep(0,5),.6,0,p,rep(0,7),p,rep(0,7),p,rep(0,7),p,rep(0,7),p,rep(0,7),.6,0,0,.4,rep(0,7),1,rep(0,9),1)
P_bb <- matrix(bb,9,9,byrow=TRUE)

(Pbb <-new('markovchain', transitionMatrix = P_bb, states = c('1','2','3','4','5','6','7','0','8'),name = 'Betting on Bail Canonical Form'))
```
##### Contruct your Q matrix
```{r}
(Q <- Pbb[1:7,1:7])
```
##### Construct the Identity matrix
```{r}
(I <- diag(dim(Q)[2]))
```
##### Calculate the Inverse martix
```{r}
(I-Q)
```
##### Claculate the fundamental matrix
```{r}
(N <- solve(I-Q))
```
##### Calcualte the number of steps before the chain is absorbed.
```{r}
c_calc <- c(rep(1,7))
c <- matrix(c_calc,7,1,byrow=TRUE)
(t <- N %*% c)
```
##### Construct the R matrix (non-zero t by r matrix)
```{r}
(R <- Pbb[1:7,8:9])
```
##### Calculate the probabilities that an absorbing chain will be absorbed
##### The probability that Smith wins 8 dollars before losing all of his money if he bets 1 dollar each time (timid strategy) is 0.0203.
```{r}
(round(B <- N %*% R,4))
```
(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy)

p(3) = 0.4p(6)

p(6) = 0.4p(8) + 0.6p(4)

p(4) = 0.4p(8)

p(0) = 0

p(8) = 1

solve and find
p(3) = 0.256, p(4) = 0.4, p(6) = 0.64

(c) By comparing the third components of the vector φ we find that the bold strategy gives Smith a better chance to get out jail

```{r}
dbinom(3,3,0.4)
```





