---
title: "Multiple linear regression"
author: ""
output:
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: inline
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, results = TRUE, fig.show = "show", message = FALSE, warning = FALSE)
```

## Grading the professor

Many college courses conclude by giving students the opportunity to evaluate 
the course and the instructor anonymously. However, the use of these student 
evaluations as an indicator of course quality and teaching effectiveness is 
often criticized because these measures may reflect the influence of 
non-teaching related characteristics, such as the physical appearance of the 
instructor. The article titled, "Beauty in the classroom: instructors' 
pulchritude and putative pedagogical productivity" by Hamermesh and Parker 
found that instructors who are viewed to be better looking receive higher 
instructional ratings. 

Here, you will analyze the data from this study in order to learn what goes 
into a positive professor evaluation.

## Getting Started

### Load packages

In this lab, you will explore and visualize the data using the **tidyverse** suite of 
packages. The data can be found in the companion package for OpenIntro resources, **openintro**.

Let's load the packages.

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(GGally)
```

This is the first time we're using the `GGally` package. You will be using the
`ggpairs` function from this package later in the lab.

### Creating a reproducible lab report

To create your new lab report, in RStudio, go to New File -> R Markdown... Then, choose From Template and then choose `Lab Report for OpenIntro Statistics Labs` from the list of templates.

### The data

The data were gathered from end of semester student evaluations for a large 
sample of professors from the University of Texas at Austin. In addition, six 
students rated the professors' physical appearance. The result is a data frame 
where each row contains a different course and columns represent variables about 
the courses and professors. It's called `evals`.

```{r}
glimpse(evals)
```

We have observations on 21 different variables, some categorical and some 
numerical. The meaning of each variable can be found by bringing up the help file:

```{r help-evals}
?evals
```

```{r}
evals
```

## Exploring the data

1.  Is this an observational study or an experiment? The original research
    question posed in the paper is whether beauty leads directly to the
    differences in course evaluations. Given the study design, is it possible to
    answer this question as it is phrased? If not, rephrase the question.
    
    Answer:  I believe this study to be experimental because the researcher is providing and
             and controlling the variables.  The researcher is not observing the subjects
             themselves.  

2.  Describe the distribution of `score`. Is the distribution skewed? What does 
    that tell you about how students rate courses? Is this what you expected to 
    see? Why, or why not?
    
    Answer:  From the histogram below the score distribution seems
             to be close to a normal distribution with a skew to the left.
             
```{r echo=TRUE}
ggplot(data = evals, aes(x = score)) +
  geom_histogram()
```

3.  Excluding `score`, select two other variables and describe their relationship 
    with each other using an appropriate visualization.

```{r echo=TRUE}
ggplot(data = evals, aes(x = age, y = bty_avg)) + geom_point()
```


```{r echo=TRUE}
ggplot(data = evals, aes(x = prof_id, y = bty_avg, color=gender)) + geom_point()
```

      Answer:  From the scatterplot above when comparing age to average beauty (bty_avg), the older an
               instructor gets the less attractive they become to their students.  This shows a negative
               linear regression relationship.  And by the second scatterplot, it appears that female 
               instructors had a higher beauty average (proportionately) than their male counterparts.

## Simple linear regression

The fundamental phenomenon suggested by the study is that better looking teachers
are evaluated more favorably. Let's create a scatterplot to see if this appears 
to be the case:

```{r scatter-score-bty_avg}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_point()
```

Before you draw conclusions about the trend, compare the number of observations 
in the data frame with the approximate number of points on the scatterplot. 
Is anything awry?

4.  Replot the scatterplot, but this time use `geom_jitter` as your layer. What 
    was misleading about the initial scatterplot?
    
    Answer:  It seems that geom_point is combining scores that are very close in 
             value thus giving it a very clean look.  While geom_jitter adds a little
             variation to each observation thus separating each point and so now each
             observation is represented in the scatterplot.

```{r scatter-score-bty_avg-jitter}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter()
```


5.  Let's see if the apparent trend in the plot is something more than
    natural variation. Fit a linear model called `m_bty` to predict average
    professor score by average beauty rating. Write out the equation for the linear 
    model and interpret the slope. Is average beauty score a statistically significant
    predictor? Does it appear to be a practically significant predictor?

    Answer:  Linear model equation:  
$\hat{y} = {\beta_{0} +\beta_{1} }$ x bty_avg

```{r m_bty}
(m_bty <- lm(score ~ bty_avg, data = evals))
```

```{r summary-m_bty}
summary(m_bty)
```

```{r}
evals %>% 
  summarise(cor(score, bty_avg))
```


      Answer: Average beauty score does not appear to be a practically significant predictor because 
              from initial model calculation the slope (bty_avg) is only 6.7% which is very low.  
              From the summary of the model m_bty, one can see the the Adjusted R-sqaured is even 
              lower at 3.3% which is a better indicator of relationship strength.

Add the line of the bet fit model to your plot using the following:
    
```{r scatter-score-bty_avg-line-se}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```

The blue line is the model. The shaded gray area around the line tells you about the 
variability you might expect in your predictions. To turn that off, use `se = FALSE`.

```{r scatter-score-bty_avg-line}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE)
```

6.  Use residual plots to evaluate whether the conditions of least squares
    regression are reasonable. Provide plots and comments for each one (see
    the Simple Regression Lab for a reminder of how to make these).
    
    Answer:  The linear regression plot with least sum of the squares line supports the
             assumption that bty_avg is not a significant predictor of score because the
             positive and negative residuals are not tightly packed along the blue line.

## Multiple linear regression

The data set contains several variables on the beauty score of the professor: 
individual ratings from each of the six students who were asked to score the 
physical appearance of the professors and the average of these six scores. Let's 
take a look at the relationship between one of these scores and the average 
beauty score.

```{r bty-rel}
ggplot(data = evals, aes(x = bty_f1lower, y = bty_avg)) +
  geom_point()

evals %>% 
  summarise(cor(bty_avg, bty_f1lower))
```

As expected, the relationship is quite strong---after all, the average score is 
calculated using the individual scores. You can actually look at the 
relationships between all beauty variables (columns 13 through 19) using the 
following command:

```{r bty-rels, echo=FALSE}
evals %>%
  select(contains("bty")) %>%
  ggpairs()
```

These variables are collinear (correlated), and adding more than one of these 
variables to the model would not add much value to the model. In this 
application and with these highly-correlated predictors, it is reasonable to use
the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score 
after you've accounted for the professor's gender, you can add the gender term 
into the model.

```{r scatter-score-bty_avg_gender}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

7.  P-values and parameter estimates should only be trusted if the
    conditions for the regression are reasonable. Verify that the conditions
    for this model are reasonable using diagnostic plots.

    Assumption 1: Are residuals nearly normal.  It appears that the residuals are not nearly
        as they seem to bow out from what would be the least sum of the squares line.
```{r}
ggplot(data = m_bty_gen, aes(sample = .resid)) +
  stat_qq()
```

    Assumption 2: The variablility of the residuals is nearly constant.  Although you can some 
        variability among the residuals there is still a lot of scattering so I would have to 
        assess that variability is not very constant.
        
```{r}
ggplot(data = m_bty_gen, aes(x = bty_avg , y = .resid)) +
  geom_point()
```
    
    Assumption 3:  I cannot definitively say that the residuals are independent because I can not 
        figure out how to plot the variables in order of their data collection
        
        
    Assumption 4:  From the obox plots and scatter plots below it appears that each varibale is 
        linearly related to the outcome.
        
```{r}
ggplot(data = m_bty_gen, aes(x = gender, y = .resid)) + geom_boxplot() + geom_point()
```

8.  Is `bty_avg` still a significant predictor of `score`? Has the addition
    of `gender` to the model changed the parameter estimate for `bty_avg`?
    
    Answer:  I still do not believe that bty_avg is a significant predictor if an
             instructor's score because all four diagnotistic assumptions are not
             true as well as the adjusted R squared value did not increase sig-
             nificantly by adding gender into the equation.

Note that the estimate for `gender` is now called `gendermale`. You'll see this 
name change whenever you introduce a categorical variable. The reason is that R 
recodes `gender` from having the values of `male` and `female` to being an 
indicator variable called `gendermale` that takes a value of $0$ for 
female professors and a value of $1$ for male professors. (Such variables 
are often referred to as "dummy" variables.)

As a result, for female professors, the parameter estimate is multiplied by zero,
leaving the intercept and slope form familiar from simple regression.

\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (0) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg\end{aligned}
\]

<!-- We can plot this line and the line corresponding to those with color pictures
with the following  -->
<!-- custom function. -->

```{r twoLines}
ggplot(data = evals, aes(x = bty_avg, y = score, color = pic_color)) +
   geom_smooth(method = "lm", formula = y ~ x, se = FALSE)
```

9.  What is the equation of the line corresponding to those with color pictures? 
    (*Hint:* For those with color pictures, the parameter estimate is multiplied
    by 1.) For two professors who received the same beauty rating, which color 
    picture tends to have the higher course evaluation score?
    
    Answer:
$\hat{y} = {\beta_{0} +\beta_{1} }$ x pic_color + $\beta_{2}$ x 0

```{r scatter-score-bty_avg_pic-color}
m_bty_pic <- lm(score ~ bty_avg + pic_color, data = evals)
summary(m_bty_pic)
```

The decision to call the indicator variable `gendermale` instead of 
`genderfemale` has no deeper meaning. R simply codes the category that 
comes first alphabetically as a $0$. (You can change the reference level of a categorical 
variable, which is the level that is coded as a 0, using the`relevel()` function. 
Use `?relevel` to learn more.)

10. Create a new model called `m_bty_rank` with `gender` removed and `rank` 
    added in. How does R appear to handle categorical variables that have more 
    than two levels? Note that the rank variable has three levels: `teaching`, 
    `tenure track`, `tenured`.
    
    Answwer:  R appears to consolidate categorical values that have a very similar 
              meaning or cannotation.  In this case, teaching and tenure track are
              not associated with tenure.

```{r scatter-score-bty_avg_rank}
m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
summary(m_bty_rank)
```

The interpretation of the coefficients in multiple regression is slightly 
different from that of simple regression. The estimate for `bty_avg` reflects
how much higher a group of professors is expected to score if they have a beauty
rating that is one point higher *while holding all other variables constant*. In
this case, that translates into considering only professors of the same rank 
with `bty_avg` scores that are one point apart.

## The search for the best model

We will start with a full model that predicts professor score based on rank, 
gender, ethnicity, language of the university where they got their degree, age, 
proportion of students that filled out evaluations, class size, course level, 
number of professors, number of credits, average beauty rating, outfit, and 
picture color.

11. Which variable would you expect to have the highest p-value in this model? 
    Why? *Hint:* Think about which variable would you expect to not have any 
    association with the professor score. 
    
    Answer: My choice would be cls_level.  Because I would think that upper level students would
            be a little harder on there instructors.  I think I believe would consider class
            level in my grading an instructor.

Let's run the model...

```{r m_full, tidy = FALSE}
m_full <- lm(score ~ rank + gender + ethnicity + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

12. Check your suspicions from the previous exercise. Include the model output
    in your response.
    
    Answer: Removing class level did not have much of an affect on the over all
            instructors score.
    
```{r m_full_no_level, tidy = FALSE}
m_full_no_level <- lm(score ~ rank + gender + ethnicity + language + age + cls_perc_eval 
             + cls_students + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full_no_level)
```

13. Interpret the coefficient associated with the ethnicity variable.

    Answer: Although it is one of the lower p-values which is a good sign, I am
        surprised that it is not lower especially with number of minority 
        students that attend UT, but, then again, I'm in Texas.

14. Drop the variable with the highest p-value and re-fit the model. Did the
    coefficients and significance of the other explanatory variables change?
    (One of the things that makes multiple regression interesting is that
    coefficient estimates depend on the other variables that are included in
    the model.) If not, what does this say about whether or not the dropped
    variable was collinear with the other explanatory variables?
    
    Answwer: Dropping the highest p-value had little affect on the adjusted R
        squared value. This tells me that the cls_profs is collinear to many
        of the other explanatory variables.
    
```{r m_full_no_cls_profs, tidy = FALSE}
m_full_no_cls_profs <- lm(score ~ rank + gender + ethnicity + language + age 
             + cls_perc_eval 
             + cls_students + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full_no_cls_profs)
```

15. Using backward-selection and p-value as the selection criterion,
    determine the best model. You do not need to show all steps in your
    answer, just the output for the final model. Also, write out the linear
    model for predicting score based on the final model you settle on.
    
```{r m_best_model, tidy = FALSE}
m_best_model <- lm(score ~ rank + gender + ethnicity + language + age 
             + cls_perc_eval + cls_students + cls_level + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_best_model)
```

16. Verify that the conditions for this model are reasonable using diagnostic 
    plots.
    
```{r}
ggplot(data = m_best_model, aes(sample = .resid)) +
  stat_qq()
```

```{r}
ggplot(data = m_bty_gen, aes(x = .resid , y = score)) +
  geom_point()
```

```{r}
ggplot(data = m_bty_gen, aes(x = score, y = .resid)) + geom_boxplot() + geom_point()
```

17. The original paper describes how these data were gathered by taking a
    sample of professors from the University of Texas at Austin and including 
    all courses that they have taught. Considering that each row represents a 
    course, could this new information have an impact on any of the conditions 
    of linear regression?
  
    Answer: If I am interpreting this correctly, it seems that some of the variables
            on the evaluation do not have a significant impact on the score given 
            to a professor.  That being said I would probably take out many of the
            variables having to do with appearance and focus more on course material,
            presentation, average grade in the class, etc.

18. Based on your final model, describe the characteristics of a professor and 
    course at University of Texas at Austin that would be associated with a high
    evaluation score.
    
    Answer: based on my final model the characteristics most important to getting 
            a positive score are rank (tenure or not), gender, language, age, number 
            of students taking the evaluation, class size, class level, number of 
            class credit, physical appearance, attire and the type of picutre posted
            (color or black and white)
             
19. Would you be comfortable generalizing your conclusions to apply to professors
    generally (at any university)? Why or why not?

## References

* * *
