---
title: "Week 9 Assignment – Web APIs"
author: "Peter Gatica"
date: "04/11/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
```{r}
library(devtools)
library(RCurl)
library(plyr)
library(ggplot2)
library(httr)
library(jsonlite)
```

## Preparing Data:
 1.  Choose one of the New York Times APIs.  I chose the Movie Reviews API 
 2.  Request API key.
 3.  Construct an interface in R to read in the JSON data that will run 3 times so that I can get 60 
     entries for my data.frame.  The max return from the API is 20.
 4.  Transformed data to an R dataframe

## Function proc_results_2_df will transform data and build the data.frame.
```{r}
proc_results_2_df <- function(results_list)
{
  (proc <- "in function")
  y <- 1
  loop_num_res = 21
  while(y < loop_num_res)
    {
    if (results_list[[y]]$display_title != "") {
    movie_title <- results_list[[y]]$display_title
    movie_rating <- results_list[[y]]$mpaa_rating
    movie_critic <- results_list[[y]]$byline
    movie_review <- results_list[[y]]$summary_short
    if (is.null(results_list[[y]]$opening_date)) {
      movie_open_date <- '9999-99-99'
    } else {
        movie_open_date <- results_list[[y]]$opening_date
    }
    if (is.null(results_list[[y]]$publication_date)) {
      movie_rev_pub_date <- '9999-99-99'
    } else {
        movie_rev_pub_date <- results_list[[y]]$publication_date
    }
    movie_review_url <- results_list[[y]]$link$url
    (new_row <- c(movie_title,movie_rating,movie_critic,movie_review,movie_open_date,movie_rev_pub_date,movie_review_url))
    movie_review_df <- rbind(movie_review_df,new_row)
    }
  if (y == 1) {
    names(movie_review_df) <- c("movie_title","movie_rating","movie_critic","movie_review","movie_open_date","movie_rev_pub_date","movie_review_url")
  }
  y <- y + 1
  }
  (df_build_status <- "movie reviews added successfully")
  return(movie_review_df)
}
```

## Initialize data frames for use
```{r}
temp_df <- data.frame(c())
movie_review_df <- data.frame(c())
movie_review_df_final <- data.frame(c())
```


## Issue API Call
1.  Issue the NYT Movie Reviews API call 3 times so that I can get 60 entries for my data.frame.
    The max return from the API is 20.
2.  Check API response status
3.  Query the NYT Movie Reviews db to get the latest 60 movie reviews 
```{r}
x <- 1
loop_num = 4
while(x < loop_num)
{
  if (x == 1) {
    url <- "https://api.nytimes.com/svc/movies/v2/reviews/picks.json?opening-date=2020-02-12:2021-04-10&order=by-opening-date&api-key=YOURKEYHERE"
    movie_reviews_1 <- GET(url)
    get_status <- movie_reviews_1$status_code
    if (get_status != 200) {
      stop_for_status(get_status)
    }
    mr_1_hdrs <- headers(movie_reviews_1)
    movie_revs_parsed_1 <- content(movie_reviews_1,"parse")
    results_list <- movie_revs_parsed_1[["results"]]
  }
  if (x == 2) {
    url <- "https://api.nytimes.com/svc/movies/v2/reviews/picks.json?opening-date=2020-01-01:2021-02-11&order=by-opening-date&api-key=YOURKEYHERE"
    movie_reviews_2 <- GET(url)
    get_status <- movie_reviews_2$status_code
    if (get_status != 200) {
      stop_for_status(get_status)
    }
    mr_2_hdrs <- headers(movie_reviews_2)
    movie_revs_parsed_2 <- content(movie_reviews_2,"parse")
    results_list <- movie_revs_parsed_2[["results"]]
  }
  if (x == 3) {
    url <- "https://api.nytimes.com/svc/movies/v2/reviews/picks.json?opening-date=2020-10-01:2020-12-03&order=by-opening-date&api-key=YOURKEYHERE"
    movie_reviews_3 <- GET(url)
    get_status <- movie_reviews_3$status_code
    if (get_status != 200) {
      stop_for_status(get_status)
    }
    mr_3_hdrs <- headers(movie_reviews_3)
    movie_revs_parsed_3 <- content(movie_reviews_3,"parse")
    results_list <- movie_revs_parsed_3[["results"]]
  }

  # Call function proc_results_2_df
  (temp_data <- proc_results_2_df(results_list))
  movie_review_df_final <- rbind(movie_review_df_final, temp_data)
  x <- x + 1
}
```

```{r}
movie_review_df_final
```
####
#### Add a unique ID column based on row name of each entry for future load to a db table.
```{r}
(movie_review_df_final <- cbind( data.frame("movie_rev_id" = as.integer(rownames(movie_review_df_final))), movie_review_df_final))
```

#### Conclusion:  I found using the NYT APIs easy to figure out.  This is one one of easier assignments for me and I think that is because it is very ETL-like.  I chose the Movie Reviews because one of my earlier assignments was building a database using MySql and I create the Movies DB and this code would fit nicely into that design.
