---
title: "Week 3 Assignment"
author: "Peter Gatica"
date: "02/19/2021"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

```{r message=FALSE}
# Load needed libraries
library(tidyverse)
library(RCurl)
library(knitr)
library(stringr)
```

### 1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r Source file from 538 website}
# Source the file from the 538 Website github repository and set NA strings to 0
filename <- getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
majors_list <- read.csv(text = filename,na.strings = "")
head(majors_list, 10)
```

```{r}
grep(pattern = 'Data', majors_list$Major, value = TRUE, ignore.case = TRUE)
```

```{r}
grep(pattern = 'statistics', majors_list$Major, value = TRUE, ignore.case = TRUE)
```

### 2. Write code that transforms the data below:

```{r}
filename <- getURL("https://raw.githubusercontent.com/audiorunner13/Masters-Coursework/main/DATA607%20Spring%202021/Week3/Data/veggies.txt")
berry_veg <- read.delim(text=filename, header = FALSE, quote = "")
berry_veg$V1
```

```{r}
berry_veg_vec <-  str_c(berry_veg,sep = ", ")
writeLines(berry_veg_vec)
```

### 3. Describe, in words, what these expressions will match

This will return an error since it is not enclosed in quotes
```{r}
# str_view(words, (.)\1\1, match = TRUE)
```

Find all words that have a pair of letters that repeat in any order
```{r}
str_view(words,"(.)(.)\\2\\1", match = TRUE)
```

This will return an error since it is not enclosed in quotes
```{r}
#str_view(words,(..)\1, match = TRUE)
```

Find all words that have a letter that repeats more than once
```{r}
str_view(words,"(.).\\1.\\1", match = TRUE)
```

Find all words that have 3 letters that repeat 0 or more times
```{r}
str_view(words,"(.)(.)(.).*\\3\\2\\1", match = TRUE)
```

### 4. Construct regular expressions to match words that:

Start and end with the same character
```{r}
words[str_detect(words, "^a.*a$")]
```

```{r}
str_view(words,"^a.*a$", match = TRUE)
```

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
```{r}
x <- c("church", "remember")
str_view(x,"(..)(..)\\1", match = TRUE)
```

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
```{r}
test_word <- c("eleven", "twelve", "twentieth","ninconponop")
str_view(test_word,"(.)(.)(.).+\\1", match = TRUE)
```
