---
title: "Week 15 Assignment"
author: "Peter Gatica"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cubature, echo = FALSE}
library(cubature)
```

WebTest Name: (Test)

1. Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.  
( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

```{r}
x = c(5.6, 6.3, 7, 7.7, 8.4)

y = c(8.8, 12.4, 14.8, 18.2, 20.8)

lm(y~x)
```
The equation for the regression line for the given points is
$$ y=4.257x\  -\  14.800 $$
2. Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form ( x, y, z ). Separate multiple points with a comma.
Using the saddle points calculator: https://calculator-online.net/saddle-point-calculator/

$$ f(x,y)=24x-6xy^{2}-8y^{3} $$
1st Derivative Steps of:
$$ \frac{d}{dx} \left( -6xy^{2}+24x-8y^{3}\right)  =24-6y^{2} $$
2nd Derivative Steps of:
$$ \frac{d}{dx} (24-6y^{2})=0 $$
1st Derivative Steps of: 
$$ \frac{d}{dx} \left( -6xy^{2}+24x-8y^{3}\right)  =-12y(x+2y) $$
2nd Derivative Steps of:
$$ \frac{d}{dx} (-12y(x+2y))=-12x-48y $$
thus $$ thus\  if\  24-6y^{2}=0\  then\  y=[2,-2] $$
Now solve for x if y=[2,-2]:
$$ f_{y}=-12xy-24y^{2}=0\  thus\  (-4,2)\  and\  (4,-2) $$

Now solve the equation:
$$ -12x-48y=-12(8)-24(-4)^{2}=0 $$
```{r}
f_eq = function(x,y){
  z = 24*x - 6*x*y^2 - 8*y^3
  return(c(x,y,z))
}
print(rbind(f_eq(-4,2),f_eq(4,-2)))
```
Answer: (-4,2,-64) and (4,2,64)

3. A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she
sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81 - 21x + 17y units of the
"house" brand and 40 + 11x - 23y units of the "name" brand.
Step 1: Find the revenue function R(x,y).
$$ R(x)=(81-21x+17y)x $$
$$ R(y)=(40+11x-23y)y $$
Combine R(x) and R(y).
$$ R(x,y)=R(x)+R(y) $$
$$ R(x,y)=28xy-23y^{2}-21x^{2}+40y+81x $$
Step 2. What is the revenue if she sells the “house” brand for \$2.30 and the “name” brand for \$4.10?

```{r}
groc_rev = function(x,y){
  g_rev = 28*x*y-23*y^2-21*x^2+40*y+81*x
  return(g_rev)
}
(revenue = groc_rev(2.3,4.1))
```
4. A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a
product each week. The total weekly cost is given by $$ C(x,y)=\frac{1}{6} x^{2}+\frac{1}{6} y^{2}+7x+25y+700 $$, x is the number of
units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in
each plant to minimize the total weekly cost?
$$ x+y=96 $$
$$ y=96-x $$
$$ x=96-y $$
Substitute the values into the equation: $$ C(x,96-x)=x^{2}+(96x)^{2}+7x+25(96-x)+700 $$
$$ C(x,96-x)=x^{2}-50x+4636 $$
Calculate the minimum value: $$ C\prime (x)=\frac{2}{3} x-50=0 $$
$$ x=75 $$
$$ 21=96-75 $$
thus Denver should produce 21 units and Los Angeles 75 to minimize total weekly costs.

5. Evaluate the double integral on the given region. $$ \iint \left( {}e^{8x+3y}\right)  dA;R:2\leq x\leq 4\  and\  2\leq y\leq 4 $$

Write your answer in exact form without decimals.  Using the Symbolad integral calculator https://www.symbolab.com/solver/double-integrals-calculator

$$ \int^{4}_{2} \int^{4}_{4} (e^{8x+3y})dxdy=$$
$$ \int^{4}_{2} (\frac{1}{8} )(e^{3y+32}-e^{3y+16})dy= $$
$$ \frac{e^{44}-e^{38}-e^{28}+e^{22}}{24}= $$
$$ \frac{{}e^{0}}{24} = 0 $$
