---
title: "The normal distribution"
author: ""
output:
  pdf_document: default
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r echo = FALSE}
knitr::opts_chunk$set(eval = TRUE, results = FALSE, fig.show = "hide", message = FALSE, warning = FALSE)
```

In this lab, you'll investigate the probability distribution that is most central
to statistics: the normal distribution.  If you are confident that your data are 
nearly normal, that opens the door to many powerful statistical methods.  Here 
we'll use the graphical tools of R to assess the normality of our data and also 
learn how to generate random numbers from a normal distribution.

## Getting Started

### Load packages

In this lab, we will explore and visualize the data using the **tidyverse** suite of packages as well as the **openintro** package.

Let's load the packages.

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Creating a reproducible lab report

To create your new lab report, in RStudio, go to New File -> R Markdown... Then, choose From Template and then choose `Lab Report for OpenIntro Statistics Labs` from the list of templates.

### The data

This week you'll be working with fast food data.  This data set contains data on
515 menu items from some of the most popular fast food restaurants worldwide.
Let's take a quick peek at the first few rows of the data.

Either you can use `glimpse` like before, or `head` to do this.

```{r load-data, results=TRUE}
library(tidyverse)
library(openintro)
head(fastfood,50)
```

You'll see that for every observation there are 17 measurements, many of which are
nutritional facts.

You'll be focusing on just three columns to get started: restaurant, calories, 
calories from fat.

Let's first focus on just products from McDonalds and Dairy Queen.

```{r load-mcdonalds, results=TRUE}
(mcdonalds <- fastfood %>%
  filter(restaurant == "Mcdonalds"))
```
```{r load-dairyqueen, results=TRUE}
(dairy_queen <- fastfood %>%
  filter(restaurant == "Dairy Queen"))
```

1.  Make a plot (or plots) to visualize the distributions of the amount of calories
from fat of the options from these two restaurants.  How do their centers, shapes,
and spreads compare?

Answer:   Both the McDonalds and Dairy Queen distribution seem very similar and appear to be skewed to the
right, McDonalds more than Dairy Queen.  Their calories from fat counts are very close to identical especially
in the 100 to 500 calorie range.  McDonalds does have a few more menu items than Dairy Queen.

```{r load-mcd-bar, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = mcdonalds, aes(x = cal_fat)) + 
  geom_bar(mapping = aes(x = cal_fat)) +
  labs(x = "McDonalds Calories from Fat", y = "Item Count")
```

```{r load-dqbar, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(x = cal_fat)) +
  geom_bar(mapping = aes(x = cal_fat)) +
  labs(x = "Dairy Queen Calories from Fat", y = "Item Count")  
```

## The normal distribution

In your description of the distributions, did you use words like *bell-shaped* 
or *normal*?  It's tempting to say so when faced with a unimodal symmetric 
distribution.

To see how accurate that description is, you can plot a normal distribution curve 
on top of a histogram to see how closely the data follow a normal distribution. 
This normal curve should have the same mean and standard deviation as the data. 
You'll be focusing on calories from fat from Dairy Queen products, so let's store
them as a separate object and then calculate some statistics that will be referenced 
later. 

```{r load-stats, results=TRUE}
(mcd.mean <- mean(mcdonalds$cal_fat))
(mcd.sd <- sd(mcdonalds$cal_fat))
(dqmean <- mean(dairy_queen$cal_fat))
(dqsd   <- sd(dairy_queen$cal_fat))
```

Next, you make a density histogram to use as the backdrop and use the `lines` 
function to overlay a normal probability curve. The difference between a 
frequency histogram and a density histogram is that while in a frequency 
histogram the *heights* of the bars add up to the total number of observations, 
in a density histogram the *areas* of the bars add up to 1. The area of each bar
can be calculated as simply the height *times* the width of the bar. Using a 
density histogram allows us to properly overlay a normal distribution curve over
the histogram since the curve is a normal probability density function that also
has area under the curve of 1. Frequency and density histograms both display the
same exact shape; they only differ in their y-axis. You can verify this by
comparing the frequency histogram you constructed earlier and the density
histogram created by the commands below.

```{r load-mcd-hist, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = mcdonalds, aes(x = cal_fat)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..)) +
        stat_function(fun = dnorm, args = c(mean = mcd.mean, sd = mcd.sd), col = "tomato") +
        labs(x = "McDonalds Calories from Fat", y = "Density")
```

```{r load-dq-hist, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(x = cal_fat)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..)) +
        stat_function(fun = dnorm, args = c(mean = dqmean, sd = dqsd), col = "tomato") +
        labs(x = "Dairy Queen Calories from Fat", y = "Density")
```

After initializing a blank plot with `geom_blank()`, the `ggplot2` package (within
the `tidyverse`) allows us to add additional layers. The first layer is a density 
histogram. The second layer is a statistical function -- the density of the normal
curve, `dnorm`. We specify that we want the curve to have the same mean and 
standard deviation as the column of female heights. The argument `col` simply sets
the color for the line to be drawn. If we left it out, the line would be drawn in
black.

2.  Based on the this plot, does it appear that the data follow a nearly normal 
    distribution?

Answer:  The Dairy Queen plot though having a slight right skew I would say that it aligns more closely to the normal
distribution curve shown.

## Evaluating the normal distribution

Eyeballing the shape of the histogram is one way to determine if the data appear
to be nearly normally distributed, but it can be frustrating to decide just how 
close the histogram is to the curve. An alternative approach involves 
constructing a normal probability plot, also called a normal Q-Q plot for 
"quantile-quantile".

```{r load-dq-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(sample = cal_fat)) + 
  geom_line(stat = "qq") +
  labs(x = "Dairy Queen theoretical")
```

This time, you can use the `geom_line()` layer, while specifying that you will 
be creating a Q-Q plot with the `stat` argument. It's important to note that here,
instead of using `x` instead `aes()`, you need to use `sample`.

The x-axis values correspond to the quantiles of a theoretically normal curve 
with mean 0 and standard deviation 1 (i.e., the standard normal distribution).  The 
y-axis values correspond to the quantiles of the original unstandardized sample 
data. However, even if we were to standardize the sample data values, the Q-Q
plot would look identical. A data set that is nearly normal will result in a
probability plot where the points closely follow a diagonal line.  Any deviations
from normality leads to deviations of these points from that line.

The plot for Dairy Queen's calories from fat shows points that tend to follow the
line but with some errant points towards the upper tail.  You're left with the same
problem that we encountered with the histogram above: how close is close enough?

A useful way to address this question is to rephrase it as: what do probability 
plots look like for data that I *know* came from a normal distribution?  We can 
answer this by simulating data from a normal distribution using `rnorm`.

```{r load-dw-sim-data, results=TRUE}
(sim_norm <- rnorm(n = nrow(dairy_queen), mean = dqmean, sd = dqsd))
```

The first argument indicates how many numbers you'd like to generate, which we 
specify to be the same number of menu items in the `dairy_queen` data set using the 
`nrow()` function.  The last two arguments determine the mean and standard 
deviation of the normal distribution from which the simulated sample will be 
generated.  You can take a look at the shape of our simulated data set, `sim_norm`, 
as well as its normal probability plot.

3.  Make a normal probability plot of `sim_norm`.  Do all of the points fall on 
    the line?  How does this plot compare to the probability plot for the real 
    data? (Since `sim_norm` is not a dataframe, it can be put directly into the 
    `sample` argument and the `data` argument can be dropped.)
    
Answer: The Dairy Queen sim_norm normal probability plat is a little smoother it is almost identical to the
actual normal plot.  Both still show that the DQ distribution for fat calories is somewhat close to normal.

```{r load-dw-simnorm-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(sample = sim_norm)) + 
  geom_line(stat = "qq") +
  labs(x = "Dairy Queen theoretical")
```

Even better than comparing the original plot to a single plot generated from a 
normal distribution is to compare it to many more plots using the following 
function. It shows the Q-Q plot corresponding to the original data in the top
left corner, and the Q-Q plots of 8 different simulated normal data.  It may be
helpful to click the zoom button in the plot window.

```{r load-dq-qqnormsim,  echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
(qqnormsim(sample = cal_fat, data = dairy_queen)) +
  labs(x = "Dairy Queen theoretical")
```

4.  Does the normal probability plot for the calories from fat look similar to the plots 
    created for the simulated data?  That is, do the plots provide evidence that the
    female heights are nearly normal?
    
Answer: Yes, the normal probability plot for the calories from fat look similar to the plots 
created for the simulated data do look almost identical.

5.  Using the same technique, determine whether or not the calories from McDonald's
    menu appear to come from a normal distribution.

Answer: Based the new plots, it is confirmed that McDonalds definitely is right skewed.  The quantile-quantile and
simulations also support that finding.

```{r load-mcd-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = mcdonalds, aes(sample = cal_fat)) + 
  geom_line(stat = "qq") +
  labs(x = "McDonalds theoretical")
```

```{r load-mcd-sim-data, results=TRUE}
(sim_norm <- rnorm(n = nrow(mcdonalds), mean = mcd.mean, sd = mcd.sd))
```

```{r load-mcd-simnorm-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = mcdonalds, aes(sample = sim_norm)) + 
  geom_line(stat = "qq") +
  labs(x = "McDonalds theoretical")
```

```{r load-mcd-qqnormsim,  echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
(qqnormsim(sample = cal_fat, data = mcdonalds)) + 
  labs(x = "McDonalds theoretical")
```

## Normal probabilities

Okay, so now you have a slew of tools to judge whether or not a variable is 
normally distributed.  Why should you care?

It turns out that statisticians know a lot about the normal distribution.  Once 
you decide that a random variable is approximately normal, you can answer all 
sorts of questions about that variable related to probability.  Take, for 
example, the question of, "What is the probability that a randomly chosen Dairy 
Queen product has more than 600 calories from fat?"

If we assume that the calories from fat from Dairy Queen's menu are normally 
distributed (a very close approximation is also okay), we can find this probability
by calculating a Z score and consulting a Z table (also called a normal probability
table).  In R, this is done in one step with the function `pnorm()`.

```{r load-dq-pnorm, results=TRUE}
1 - pnorm(q = 600, mean = dqmean, sd = dqsd)
```

Note that the function `pnorm()` gives the area under the normal curve below a 
given value, `q`, with a given mean and standard deviation.  Since we're 
interested in the probability that a Dairy Queen item has more than 600 calories
from fat, we have to take one minus that probability.

Assuming a normal distribution has allowed us to calculate a theoretical 
probability.  If we want to calculate the probability empirically, we simply 
need to determine how many observations fall above 600 then divide this number 
by the total sample size.

```{r load-dq-summarise, results=TRUE}
dairy_queen %>% 
  filter(cal_fat > 600) %>%
  summarise(percent = n() / nrow(dairy_queen))
```

Although the probabilities are not exactly the same, they are reasonably close. 
The closer that your distribution is to being normal, the more accurate the 
theoretical probabilities will be.

6.  Write out two probability questions that you would like to answer about any 
    of the restaurants in this dataset.  Calculate those probabilities using both
    the theoretical normal distribution as well as the empirical distribution 
    (four probabilities in all).  Which one had a closer agreement between the 
    two methods?
    
6a. What is the probability that a randomly chosen Dairy Queen product has more than 40 grams of protein? 

Answer:
Caculate the mean and standard deviation of protein in DQ menu items;
```{r load-dq-protein-mean-sd, results=TRUE}
(dqpmean <- mean(dairy_queen$protein))
(dqpsd   <- sd(dairy_queen$protein))
```

```{r load-dqpbar, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(x = protein)) +
  geom_bar(mapping = aes(x = protein)) +
  labs(x = "Dairy Queen Protein", y = "Item Count")  
```

```{r load-dqp-hist, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(x = protein)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..)) +
        stat_function(fun = dnorm, args = c(mean = dqpmean, sd = dqpsd), col = "tomato") +
        labs(x = "Dairy Queen Protein", y = "Density")
```

Construct a normal probability Q-Q "quantile-quantile" plot for protein.

```{r load-dq-prtn-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(sample = protein)) + 
  geom_line(stat = "qq") +
  labs(x = "Dairy Queen theoretical")
```

```{r load-dq-prtn-qqnormsim,  echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
(qqnormsim(sample = protein, data = dairy_queen)) + 
  labs(x = "Dairy Queen theoretical")
```

The normal probability plot for the grams of protein in a menu item look similar to the plots created for the simulated data very close to identical.  This shows that the DQ distribution for protein content is somewhat close to normal.

Theoretically, there is a 9.5% probability that I can randomly select a menu item that has more than 40g of protein.
```{r load-dq-prtn-pnorm, results=TRUE}
1 - pnorm(q = 40, mean = dqpmean, sd = dqpsd)
```
Empirically, there is a 11.9% chance that I can randomly select a menu item that has more than 40g of proten.
```{r load-dqp-summarise, results=TRUE}
dairy_queen %>% 
  filter(protein > 40) %>%
  summarise(percent = n() / nrow(dairy_queen))
```

6b. What is the probability that a randomly chosen Dairy Queen product has more than 40 grams of carbs? 

Answer:
Calculate the mean and standard deviation of protein in DQ menu items;
```{r load-dq-totcarb-mean-sd, results=TRUE}
(dqcbmean <- mean(dairy_queen$total_carb))
(dqcbsd   <- sd(dairy_queen$total_carb))
```

```{r load-dqcbbar, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(x = total_carb)) +
  geom_bar(mapping = aes(x = total_carb)) +
  labs(x = "Dairy Queen Total Carbs", y = "Item Count")  
```

```{r load-dqcb-hist, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(x = total_carb)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..)) +
        stat_function(fun = dnorm, args = c(mean = dqcbmean, sd = dqcbsd), col = "tomato") +
        labs(x = "Dairy Queen Total Carbs", y = "Density")
```

Construct a normal probability Q-Q "quantile-quantile" plot for total carbs

```{r load-dq-tcb-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(sample = total_carb)) + 
  geom_line(stat = "qq") +
  labs(x = "Dairy Queen theoretical")
```

```{r load-dq-tcb-qqnormsim,  echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
(qqnormsim(sample = total_carb, data = dairy_queen)) + 
  labs(x = "Dairy Queen theoretical")
```

The normal probability plot for the grams of protein in a menu item look similar to the plots created for the simulated data very close to identical.  This shows that the DQ distribution for protein content is somewhat close to normal.

Theoretically, there is a 39.5% probability that I can randomly select a menu item that has more than 45g total carbs.
```{r load-dq-tcb-pnorm, results=TRUE}
1 - pnorm(q = 45, mean = dqcbmean, sd = dqcbsd)
```
Empirically, there is a 21.43% chance that I can randomly select a menu item that has more than 40g of proten.
```{r load-dqcb-summarise, results=TRUE}
dairy_queen %>% 
  filter(total_carb > 45) %>%
  summarise(percent = n() / nrow(dairy_queen))
```

* * *

## More Practice

7.  Now let's consider some of the other variables in the dataset.  Out of all the
    different restaurants, which ones' distribution is the closest to normal for sodium?
    
Answer:  Burger King and Subway seem to have the closest distribution to normal for sodium of all the restaurants in
the dataset.
    
```{r load-dq-sdm-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = dairy_queen, aes(sample = sodium)) + 
  geom_line(stat = "qq") +
  labs(x = "Dairy Queen Sodium theoretical")
```

```{r load-mcd-sodm-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
ggplot(data = mcdonalds, aes(sample = sodium)) + 
  geom_line(stat = "qq") +
  labs(x = "McDonalds sodium theoretical")
```

```{r load-dq-cf-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
cf <- fastfood %>%
  filter(restaurant == "Chick Fil-A")

ggplot(data = cf, aes(sample = sodium)) + 
  geom_line(stat = "qq") +
  labs(x = "Chick Fil-A Sodium theoretical")
```

```{r load-dq-sn-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
sonic <- fastfood %>%
  filter(restaurant == "Sonic")

ggplot(data = sonic, aes(sample = sodium)) + 
  geom_line(stat = "qq") +
  labs(x = "Sonic Sodium theoretical")
```

```{r load-dq-arb-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
arb <- fastfood %>%
  filter(restaurant == "Arbys")

ggplot(data = cf, aes(sample = sodium)) + 
  geom_line(stat = "qq") +
  labs(x = "Arbys Sodium theoretical")
```

```{r load-dq-bk-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
bk <- fastfood %>%
  filter(restaurant == "Burger King")

ggplot(data = bk, aes(sample = sodium)) + 
  geom_line(stat = "qq") +
  labs(x = "Burger King Sodium theoretical")
```
```{r load-dq-sw-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
sw <- fastfood %>%
  filter(restaurant == "Subway")

ggplot(data = sw, aes(sample = sodium)) + 
  geom_line(stat = "qq") +
  labs(x = "Subway Sodium theoretical")
```
```{r load-dq-tb-qq, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
tb <- fastfood %>%
  filter(restaurant == "Taco Bell")

ggplot(data = tb, aes(sample = sodium)) + 
  geom_line(stat = "qq") +
  labs(x = "Taco Bell Sodium theoretical")
```

8.  Note that some of the normal probability plots for sodium distributions seem
    to have a stepwise pattern. why do you think this might be the case?

9.  As you can see, normal probability plots can be used both to assess 
    normality and visualize skewness.  Make a normal probability plot for the total
    carbohydrates from a restaurant of your choice.  Based on this normal 
    probability plot, is this variable left skewed, symmetric, or right skewed?  
    Use a histogram to confirm your findings.

Answer:  See 6B.  The histogram for total carbs for Dairy Queen is slightly right skewed.


* * *
