---
title: "Untitled"
author: "Connolly, Gatica, and Reeves"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### The Premise

We want to predict the performance of Ethereum against USD using historical data.  Generally, this practice is frowned upon because each new trading day changes the business cycles that produce following rates.  However, for the sake of this assignment, instead of using regressors, we are going to try to create a model based on categorization and feature engineering.

Our data is a bit rough, with only daily readings of Open, High, Low, Close, Volume, (OHLC) and Block Size.  Finer-grain data is a bit expensive.

Instead of predicting an exact exchange rate, we will simply try to determine if the following week from any data point is bullish (significantly increasing) or bearish (significantly decreasing).

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(zoo)
library(smooth)
library(TTR)
library(tidymodels)
library(ranger)
library(plotly)
library(quantmod)
data <- "https://raw.githubusercontent.com/TheWerefriend/data607Final/main/ETH-USD.csv" %>%
  read.csv(header = TRUE) %>%
  tibble() %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))
data <- data[, -6] %>%
  mutate(Open = as.numeric(as.character(Open))) %>%
  mutate(High = as.numeric(as.character(High))) %>%
  mutate(Low = as.numeric(as.character(Low))) %>%
  mutate(Close = as.numeric(as.character(Close))) %>%
  mutate(Volume = as.numeric(as.character(Volume))) %>%
  na.locf()
# na.locf() from zoo replaces NA with previous non-NA value
```

### Feature engineering

We will be using volume to establish a feature called Momentum, which is the rate of change of the volume of trades.  This will tell us if there is significant investor emotion regarding this asset.

```{r}
period <- 9
data <- data %>%
  arrange(Date) %>%
  mutate(Momentum = Volume - lag(Volume,
                                 n = period,
                                 order_by = Date))
head(data,20)
```

The next feature will be MACD (Moving Average Convergence/Divergence), which shows differences between exponential moving averages of two different periods.  If momentum is significant, this feature will tell us what direction the price is moving.

```{r}
fast <- 12
slow <- 26
data <- data %>%
  arrange(Date) %>%
  mutate(SMA12 = sma(Close, h = fast)$fitted[,1])
MACD <- MACD(data$Close, nFast = fast, nSlow = slow)
data <- cbind(data, MACD)
```

Next, we will include a Stochastic oscillator, which gives the relationship between the asset's closing price and the price range over the previous period.

```{r}
stochastic <- stoch(data$Close, nFastK = 14, nFastD = 3, nSlowD = 3)
data <- cbind(data, stochastic)
```

Finally, we must look into the future to score the Outlook for each observation as bullish or not.

```{r}
data <- data %>%
  arrange(Date) %>%
  mutate(Future = as.factor(ifelse(lead(Close, n = 7) - Close > 0, TRUE, FALSE)))
```

We now must remove the trailing NAs we introduced with out lagged/leading indicators.

```{r}
data <- na.omit(data)
```

### The Model

We will split the data at 2/3 and train a model (probably random forest) with independent variables Momentum, MACD, Stochastic, and dependent being the Outlook.

```{r}
split <- initial_split(data, strata = Future, p = 0.67)
trainer <- training(split)
tester <- testing(split)
ethRecipe <- recipe(Future ~ Open + High + Low + Close +
                      Volume + Momentum + SMA12 + macd +
                      signal + fastK + fastD + slowD,
                    data = trainer) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors()) %>%
  prep(training = trainer)
ethTrain <- juice(ethRecipe)
ethTest <- bake(ethRecipe, tester)
ethModel <- rand_forest(mode = "classification") %>%
  set_engine("ranger")
ethFit <- ethModel %>%
  fit(Future ~ Open + High + Low + Close + Volume + 
        Momentum + SMA12 + macd + signal + fastK + fastD + slowD,
        data = tester)
```

### Scoring the Model

Precision? Recall? F-measure?

```{r}
results <- ethTest %>%
  select(Future) %>%
  mutate(predicted = factor(ifelse(ethFit$fit$predictions[,1] > 0.5, TRUE, FALSE)))
precision(results, truth = Future, estimate = predicted)
recall(results, truth = Future, estimate = predicted)
f_meas(results, truth = Future, estimate = predicted)
```

### Visualizations

Use the levels function to determine the positive and negative classes.  Here we see that FALSE is the positive class.

```{r}
levels(data[["Future"]])
```

Confusion matrix

Confusion Matrix - Our model correctly classified 220 out of 332 predictions (66.3%).  Which translates into:
  - 126 correct FALSE predictions
  - 94  correct True predications

```{r}
conf_mat(results,
         truth = Future,
         estimate = predicted)
```

Line plots? Box plots?

By using plotly, an interactice plotting library, it's possible to chart financial data such as ours. 

## Candlestick Chart
Source: https://plotly.com/r/candlestick-charts/

```{r}
data_cs <- as.data.frame(data)

#cs <- data_cs %>% plot_ly(x = ~Date, type = "candlestick",
#                          open = ~Open, close = ~Close,
#                          high = ~High, low = ~Low)
#cs <- cs %>% layout(title = "Near all-time $Eth Candlestick Chart")
#cs
```

```{r}
data_cs <- data.frame(data_idx=index(data_cs),coredata(data_cs))
```

```{r}
# create Bollinger Bands
bbands <- BBands(data_cs[,c("High","Low","Close")])
```

```{r}
# join and subset data
data_cs <- subset(cbind(data_cs, data.frame(bbands[,1:3])), Date >= "2015-08-06")
```

```{r}
# colors column for increasing and decreasing
for (i in 1:length(data_cs[,1])) {
  if (data_cs$Close[i] >= data_cs$Open[i]) {
      data_cs$direction[i] = 'Increasing'
  } else {
      data_cs$direction[i] = 'Decreasing'
  }
}
```

```{r}
increase_color <- list(line = list(color = '#17BECF'))
decrease_color <- list(line = list(color = '#7F7F7F'))
```

```{r}
cs_chart <- data_cs %>% plot_ly(x = ~Date, type="candlestick",
          open = ~Open, close = ~Close,
          high = ~High, low = ~Low, name = "ETH",
          increasing = increase_color, decreasing = decrease_color)

cs_chart <- cs_chart %>% add_lines(x = ~Date, y = ~dn, name = "B Bands",
            line = list(color = '#ccc', width = 0.5),
            legendgroup = "Bollinger Bands", inherit = F,
            showlegend = FALSE, hoverinfo = "none") 

cs_chart <- cs_chart %>% add_lines(x = ~Date, y = ~mavg, name = "Mv Avg",
            line = list(color = '#E377C2', width = 0.5),
            hoverinfo = "none", inherit = F)

cs_chart <- cs_chart %>% layout(yaxis = list(title = "Price"))
```

```{r}
# plot volume bar chart
cs_chart_2 <- data_cs 
cs_chart_2 <- cs_chart_2 %>% plot_ly(x=~Date, y=~Volume, type='bar', name = "ETH Volume",
          color = ~direction, colors = c('#17BECF','#7F7F7F')) 
cs_chart_2 <- cs_chart_2 %>% layout(yaxis = list(title = "Volume"))
```

```{r}
# create rangeselector buttons
range_sel <- list(visible = TRUE, x = 0.5, y = -0.055,
           xanchor = 'center', yref = 'paper',
           font = list(size = 9),
           buttons = list(
             list(count=1,
                  label='RESET',
                  step='all'),
             list(count=1,
                  label='1 YR',
                  step='year',
                  stepmode='backward'),
             list(count=3,
                  label='3 MO',
                  step='month',
                  stepmode='backward'),
             list(count=1,
                  label='1 MO',
                  step='month',
                  stepmode='backward')
           ))
```

```{r}
# subplot with shared x axis
cs_chart <- subplot(cs_chart, cs_chart_2, heights = c(0.7,0.2), nrows=2,
             shareX = TRUE, titleY = TRUE)
cs_chart <- cs_chart %>% layout(title = paste("ETH: 2015-09-01 -",Sys.Date()),
         xaxis = list(rangeselector = range_sel),
         legend = list(orientation = 'h', x = 0.5, y = 1,
                       xanchor = 'center', yref = 'paper',
                       font = list(size = 10),
                       bgcolor = 'transparent'))
```

```{r}
cs_chart
```

### Conclusions

This model is absolute crap.