---
title: "Final Project - Etherium Cryptocurrency"
subtitle: "DATA607 - Acquisition of Data and Management - Instructor: Andrew Catlin"
author: "Authors: Joe Connelly, Peter Gatica"
date: "Date: 5/19/2021"
output: html_notebook
---

```{r message=FALSE}
# Load needed libraries
library(devtools)
library(tidyverse)
library(RCurl)
library(plyr)
library(knitr)
library(RMySQL)
```

Function call to MySQL db to connect and use the return command in a function.

```{r MySQL Connect Function}
conn.MyQL <- function(db_parms)
{
  db_conn <- dbConnect(MySQL(), user=db_user, password=db_password, dbname=db_name, host=db_host)
  return(db_conn)
}
```

Source the credentials parameter file from your local directory.  Do not store in github repository since R has no encryption capability.

```{r Read the user MySQl Parameter credentials file}
filename <- "Your Credentials Go Here"
db_parms <- read.csv(filename)
```

Set the login application credential to pass to the db log in function

```{r Set  up DB parms}
db_user = db_parms$db_user
db_password = db_parms$db_password
db_name = db_parms$db_name
db_host = db_parms$db_host
db_result_set = ""

db_parms <- c(db_user, db_password, db_name, db_host, db_result_set)
```

Call the conn.MySql connect function to access the movies database

```{r Connect to the MySQL db}
db_conn <- conn.MyQL(db_parms)
```

Use the dbListTables() function to list the tables in the database

```{r List the tables in the movies db}
dbListTables(db_conn)
```

Use the dbListFields() function to list the fields in a database table

```{r List the fields in the movie dimension database table}
dbListFields(db_conn, "ether_block_size")
```

Source etherium block size file from the ether github repository to load to the movie dimension

```{r Source the movie dimension file to load}
filename <- getURL("https://raw.githubusercontent.com/audiorunner13/Masters-Coursework/main/DATA607%20Spring%202021/Final%20Project/Data/export-BlockSize.csv")
eth_blk_size_df <- read.csv(text=filename)
```

```{r}
names(eth_blk_size_df) <- c("ether_dt", "ether_ts", "ether_blk_size")
eth_blk_size_df$ether_dt <- as.Date(eth_blk_size_df$ether_dt, format = "%m/%d/%Y")
```

```{r}
head(eth_blk_size_df,10)
```

Drop tables if they exist. Dropping and reloading is only recommended when table size and contents is small.  Write the database tables from their respective data frames.

```{r Drop table and reload the movie dimension}
#if (dbExistsTable(db_conn, "ether_block_size"))
#    dbRemoveTable(db_conn, "ether_block_size")

dbSendQuery(db_conn, "SET GLOBAL local_infile = true;")
dbWriteTable(db_conn, name = "ether_block_size", value = eth_blk_size_df, append=TRUE,row.names = FALSE)
```

Source the customer dimension file from the movies github repository to load to the customer dimension

```{r Source the customer dimension file to load}
filename <- getURL("https://raw.githubusercontent.com/audiorunner13/Masters-Coursework/main/DATA607%20Spring%202021/Final%20Project/Data/export-EtherPrice.csv")
eth_price_df <- read.csv(text=filename)
```

```{r}
head(eth_price_df,10)
```

```{r}
names(eth_price_df) <- c("ether_dt", "ether_ts", "ether_price")
eth_price_df$ether_dt <- as.Date(eth_price_df$ether_dt, format = "%m/%d/%y")
```

```{r Drop table and reload for customer dimension}
#if (dbExistsTable(db_conn, "ether_price_tbl"))
#    dbRemoveTable(db_conn, "ether_price_tbl")

dbSendQuery(db_conn, "SET GLOBAL local_infile = true;")
dbWriteTable(db_conn, name = "ether_price_tbl", value = eth_price_df, append=TRUE, row.names = FALSE)
```

Source the movie rank survey results file from the movies github repository.

```{r Source the movie rank survey file to load}
filename <- getURL("https://raw.githubusercontent.com/audiorunner13/Masters-Coursework/main/DATA607%20Spring%202021/Final%20Project/Data/export-TxGrowth.csv")
ether_trans_df <- read.csv(text=filename)
```

```{r}
names(ether_trans_df) <- c("ether_dt", "ether_ts", "ether_cnt")
ether_trans_df$ether_dt <- as.Date(ether_trans_df$ether_dt, format = "%m/%d/%Y")
```

```{r Drop table and reload for small fact table}
#if (dbExistsTable(db_conn, "ether_trans"))
#    dbRemoveTable(db_conn, "ether_trans")

dbSendQuery(db_conn, "SET GLOBAL local_infile = true;")
dbWriteTable(db_conn, name = "ether_trans", value = ether_trans_df, append=TRUE, row.names = FALSE)
```

Source the sql file in the ether github repositpry.  The sql will extract all survey answers and order by them first name, last name, and movie title and will replace nulls in the AFI 100 Rank field if a movie is not ranked.

```{r source sql file and substitute each new line "\n" with a space}
filename <- "/Users/Audiorunner13/CUNY MSDS Course Work/DATA607 Spring 2021/Final Project/SQL/ether_extract.sql"
db_sql <- readChar(filename, file.info(filename)$size)
db_sql <- gsub("\n", " ",db_sql)
db_sql
```

Execute the sql query joining the fact table to the dimension tables and return all records in the result set.  Specify the number of records to return by adjusting the "n = " argument.

```{r execute sql query and return result set}
db_data = dbSendQuery(db_conn, db_sql)
result_set = fetch(db_data, n = -1)
head(result_set,10)
```