---
title: "Project 1 - Chess Tournament Rankings"
author: "Peter Gatica"
date: "02/27/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load needed libraries
```{r message=FALSE}
# Load needed libraries
library(devtools)
library(tidyverse)
library(RCurl)
library(plyr)
library(knitr)
```

Source the tournamentinfo.txt.  Although I could not figure out how to process this messy file, after so many hours trying I decided to run a few sed commandrs like sed -e 's/\r$//', which get rid of carriage returns.  I also cleaned up the file by using a couple of sustitution commands while editing the file.
```{r}
filename <- getURL("https://raw.githubusercontent.com/audiorunner13/Masters-Coursework/main/DATA607%20Spring%202021/Week%204/Project1/Data/tournamentinfo_unix.txt")
chess_trny <- read.delim(text=filename,header=TRUE, sep = "|")
chess_trny
```

select data for the columns from the data file below:
```{r}
chess_scores <- chess_trny[, c("PairNum", "PlayerName", "Total","Round","Round.1","Round.2","Round.3","Round.4","Round.5","Round.6","State","USCF.ID...Rtg..Pre.Post.","N")]
chess_scores
```

Rename the column names
```{r}
names(chess_scores) <- c("Pair_Num","Player_Name","Total_Points","Round_1","Round_2","Round_3","Round_4","Round_5","Round_6","Round_7","State","USCF_ID_Rtg_Pre_Post","N")
chess_scores
```

Prep initial out file
```{r}
chess_results <- chess_scores[,c("Pair_Num","Player_Name","State","Total_Points")]
chess_results
```

create a vector player rankings
```{r}
ranking <- as.numeric(str_sub(chess_scores$USCF_ID_Rtg_Pre_Post, 15, 19))
ranking
```

Add the ranking(rating) to the our data.frame
```{r}
chess_results1 <- chess_results
chess_results1$Rating <- ranking
chess_results1
```

Function to calculate the opponents' average chess ranking
```{r Calculate Avg Rankin}
calc.avg.rtg <- function(loop_num)
{
  avg.rtg <- 0
  tot_player_opp_rtg <- 0
  i = 1
  for (i in 1:7) {
     if (i==1) {
       game <- chess_scores[chess_scores$Pair_Num == loop_num,  c("Round_1")]
       game_result <- str_sub(game, 1, 1)
       if (game_result == 'W' | game_result == "L" | game_result == "D") {
          player_opp <- str_sub(game, 4,5)
          player_opp_rtg <- chess_results1[chess_results1$Pair_Num == as.numeric(player_opp),  c("Rating")]
          tot_player_opp_rtg = tot_player_opp_rtg + player_opp_rtg
          player_opp_rtg = 0
       }
     }
     if (i==2){
           game <- chess_scores[chess_scores$Pair_Num == loop_num,  c("Round_2")]
           game_result <- str_sub(game, 1, 1)
           if (game_result == 'W' | game_result == "L" | game_result == "D") {
             player_opp <- str_sub(game, 4,5)
             player_opp_rtg <- chess_results1[chess_results1$Pair_Num == as.numeric(player_opp),  c("Rating")]
             tot_player_opp_rtg = tot_player_opp_rtg + player_opp_rtg
             player_opp_rtg = 0
           }
     }
     if (i==3){
           game <- chess_scores[chess_scores$Pair_Num == loop_num,  c("Round_3")]
           game_result <- str_sub(game, 1, 1)
           if (game_result == 'W' | game_result == "L" | game_result == "D") {
             player_opp <- str_sub(game, 4,5)
             player_opp_rtg <- chess_results1[chess_results1$Pair_Num == as.numeric(player_opp),  c("Rating")]
             tot_player_opp_rtg = tot_player_opp_rtg + player_opp_rtg
             player_opp_rtg = 0
           }
     }
     if (i==4){
        game <- chess_scores[chess_scores$Pair_Num == loop_num,  c("Round_4")]
        game_result <- str_sub(game, 1, 1)
        if (game_result == 'W' | game_result == "L" | game_result == "D") {
          player_opp <- str_sub(game, 4,5)
          player_opp_rtg <- chess_results1[chess_results1$Pair_Num == as.numeric(player_opp),  c("Rating")]
          tot_player_opp_rtg = tot_player_opp_rtg + player_opp_rtg
          player_opp_rtg = 0
        }
     }
     if (i==5){
       game <- chess_scores[chess_scores$Pair_Num == loop_num,  c("Round_5")]
       game_result <- str_sub(game, 1, 1)
       if (game_result == 'W' | game_result == "L" | game_result == "D") {
         player_opp <- str_sub(game, 4,5)
         player_opp_rtg <- chess_results1[chess_results1$Pair_Num == as.numeric(player_opp),  c("Rating")]
         tot_player_opp_rtg = tot_player_opp_rtg + player_opp_rtg
         player_opp_rtg = 0
       }
     }
     if (i==6){
       game <- chess_scores[chess_scores$Pair_Num == loop_num,  c("Round_6")]
       game_result <- str_sub(game, 1, 1)
       if (game_result == 'W' | game_result == "L" | game_result == "D") {
         player_opp <- str_sub(game, 4,5)
         player_opp_rtg <- chess_results1[chess_results1$Pair_Num == as.numeric(player_opp),  c("Rating")]
         tot_player_opp_rtg = tot_player_opp_rtg + player_opp_rtg
         player_opp_rtg = 0
       }
     }
     if (i==7){
       game <- chess_scores[chess_scores$Pair_Num == loop_num,  c("Round_7")]
       game_result <- str_sub(game, 1, 1)
       if (game_result == 'W' | game_result == "L" | game_result == "D") {
         player_opp <- str_sub(game, 4,5)
         player_opp_rtg <- chess_results1[chess_results1$Pair_Num == as.numeric(player_opp),  c("Rating")]
         tot_player_opp_rtg = tot_player_opp_rtg + player_opp_rtg
         player_opp_rtg = 0
       }
     }
  }  
  avg.rtg <- tot_player_opp_rtg / 7
  return(avg.rtg)
}
```

Call the function to calculate the opponents' average chess ranking
```{r}
avg_rate_all_vec <- c()
loop_num <- nrow(chess_scores) + 1
x <- 1
while(x < loop_num)
{
  avg_rate <- calc.avg.rtg(x)
  avg_rate_all_vec <- rbind(avg_rate_all_vec, avg_rate)
  x <- x + 1
}
```


Add the opponents average ranking
```{r}
chess_results1_fin <- chess_results1
chess_results1_fin$Avg_Opp_Rank <- round(avg_rate_all_vec,0)
chess_results1_fin
```

Write output file to a csv file.
```{r Write a csv file to your directory}
out_file <- "/Users/Audiorunner13/CUNY MSDS Course Work/DATA607 Spring 2021/Week4/Project1/Data/chess_results.csv"
write_csv(chess_results1_fin,out_file)
```

