---
title: "DATA 606 Data Project Proposal"
author: "Peter Gatica"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

```{r load-packages, message=TRUE, include=FALSE}
library(DATA606)
library(tidyverse)
library(openintro)
library(ggplot2)
```

### Data Preparation

The data I'm working with is in the openintro package and it's called `hfi`, 
short for Human Freedom Index.

```{r}
### The data

data("hfi")
```

```{r}
head(hfi,10)
```

There are a lot of columns in this data. I will take a subset the data to include only the variables I am interested in. I will also rename the columns to be more descriptive.

```{r}
(hfI_subset <- hfi[,c("year","ISO_code","countries","region","pf_rol","pf_religion_restrictions","pf_religion","pf_expression_control","pf_expression","pf_score","pf_rank","ef_money_inflation","ef_money_currency","ef_money","ef_score","ef_rank","hf_score","hf_rank","hf_quartile")])
```

```{r include=FALSE}
(hfi_west <- hfI_subset %>%
   filter(year == "2016") %>%
   filter(region == "North America" | region == "Western Europe"))
```

```{r}
arrange(hfi_west, desc(region))
```

```{r include=FALSE}
(hfI_latin_amer <- hfI_subset %>%
   filter(year == "2016") %>%
   filter(region == "Latin America & the Caribbean"))
```

```{r}
arrange(hfI_latin_amer, desc(region))
```

### Research question 

Does the happiness factor differ from world region by region based on religious freedom, economic freedom and freedom of expression?

### Cases 

Each case represents a region of the world such as Western Europe and N. America, Eastern Europe, etc.

### Data collection 

The report is co‐published by the [Cato Institute](https://www.cato.org/human-freedom-index/2020) and the [Fraser Institute](https://www.fraserinstitute.org/).

### Type of study 

This is an observational study.

### Data Source 

The [Human Freedom Index](https://www.cato.org/human-freedom-index/2020) presents the state of human freedom in the world based on a broad measure that encompasses personal, civil, and economic freedom. Human freedom is a social concept that recognizes the dignity of individuals and is defined here as negative liberty or the absence of coercive constraint. Because freedom is inherently valuable and plays a role in human progress, it is worth measuring carefully. The Human Freedom Index is a resource that can help to more objectively observe relationships between freedom and other social and economic phenomena, as well as the ways in which the various dimensions of freedom interact with one another.

The report is co‐published by the [Cato Institute](https://www.cato.org/human-freedom-index/2020) and the [Fraser Institute](https://www.fraserinstitute.org/).

### Response 

The response variable is 2016 happiness factor and is numerical.


### Explanatory 

The explanatory variables are ef_score and pf_score and are numerical.

### Relevant summary statistics 

**Provide summary statistics relevant to your research question. For example, if you’re comparing means across groups provide means, SDs, sample sizes of each group. This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
ggplot(data = hfi_west, aes(x = pf_expression_control, y = hf_score)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```

```{r}
ggplot(data = hfI_latin_amer, aes(x = pf_religion, y = pf_score)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```