---
title: "DATA 606 Data Project"
author: "Peter Gatica"
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r load-packages, message=TRUE, include=FALSE}
library(DATA606)
library(tidyverse)
library(openintro)
library(ggplot2)
```


## Part 1 - Introduction
### Abstract

##### The purpose of this project is to explore the possible reasons that citizens in countries around the world consume alcohol.  
After reviewing the alcohol consumption dataset, I did not see a meaningful way on how

For this project I download both the datasets from FiveThirtyEight and the CATO Institute for 2010.  I then manually combined variables from both datasets to use for my final project research.  I matched the data from both sources by country, taking person freedom, economic freedom, and the overall happiness factor from each country and combined them with the 

## Part 2 - Data

### Data Sources 

##### [FiveThirtyEight](https://github.com/fivethirtyeight/data/tree/master/alcohol-consumption) - The dataset that I am using as one of my sources was found that the FiveThirtyEight github link.  This dataset on alcohol consumption by country for 2010 is the data behind the article [Dear Mona Followup:Where Do People Drink The Most Beer, Wine and Spirits](https://fivethirtyeight.com/features/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits/).  The data was collected by the [World Health Organization](https://www.who.int/)

##### The [Human Freedom Index](https://www.cato.org/human-freedom-index/2020) presents the state of human freedom in the world based on a broad measure that encompasses personal, civil, and economic freedom. Human freedom is a social concept that recognizes the dignity of individuals and is defined here as negative liberty or the absence of coercive constraint. Because freedom is inherently valuable and plays a role in human progress, it is worth measuring carefully. The Human Freedom Index is a resource that can help to more objectively observe relationships between freedom and other social and economic phenomena, as well as the ways in which the various dimensions of freedom interact with one another.

##### The report is co‐published by the [Cato Institute](https://www.cato.org/human-freedom-index/2020) and the [Fraser Institute](https://www.fraserinstitute.org/).

### Data collection 

```{r}
filename <- "/Users/Audiorunner13/CUNY MSDS Course Work/DATA606 Spring 2021/Final Project/Data/alc_hfi_2010_data.csv"
alc_hfi_2010 <- read.csv(filename)
```

```{r}
(alc_hfi_2010 <- na.omit(alc_hfi_2010))
```
### Description of the dependent variable (what is being measured?)

The response variable isamount of alcohol (litres) consumed per person by country in 2010 and is numerical.

### Explanatory 

### Description of the independent variable (what is being measured?, include at least 2 variables)

The explanatory variables are happiness score (hf_score), personal freedom score (pf_score), and economic freedom score (ef_score).
Personal freedom and economic scores contribute to the overall happiness score of a country's citizens.

## Research question 

Does the happiness factor score of a country's citizens affected the amount of alcohol consumed by that country?

### Cases 

Each case represents the overall human factor score of each country


### Type of study 

This is an observational study.

### Research Question

### Summary Statistics
```{r}
summary(alc_hfi_2010)
```

## Part 3 - Exploratory data analysis

### Appropriate Visualizations

#bar graph of top 10 soft skills as percentage
```{r plot bar graph of top 15 alc consume, echo=TRUE, fig.keep='all', message=FALSE, results='hide'}
head(alc_hfi_2010[order(-alc_hfi_2010$total_litres_of_pure_alcohol),],15) %>% 
    ggplot(aes(y=reorder(countries,total_litres_of_pure_alcohol),x=total_litres_of_pure_alcohol,fill=countries)) +
      geom_bar(stat = 'identity',position=position_dodge()) +
      geom_text(aes(label=total_litres_of_pure_alcohol), vjust=1.0, color="black", 
            position = position_dodge(0.9), size=3.0) +
      labs(x = ("Total Alcohol Servings"),y = ("Country"),
      title = ("Top 15 Countries in Alcohol Consumed (litres) in 2010 per Person")  )
      theme_minimal()
```

```{r}
# head(arrange(alc_hfi_2010),15)
alc_hfi_2010[order(alc_hfi_2010$hf_rank),]
```

#bar graph of top 10 soft skills as percentage
```{r echo=TRUE, fig.keep='all', message=FALSE, results='hide'}
head(alc_hfi_2010[order(-alc_hfi_2010$hf_score),],15) %>% 
    ggplot(aes(y=reorder(countries,hf_score),x=hf_score,fill=countries)) +
      geom_bar(stat = 'identity',position=position_dodge()) +
      geom_text(aes(label=hf_score), vjust=1.0, color="black", 
            position = position_dodge(0.9), size=3.0) +
      labs(x = ("Happiness Factor Rank"),y = ("Country"),
      title = ("Top 15 Countries Happiness Factor Score")  )
      theme_minimal()
```

### Statistical Output
### For regression models, include the regression output and interpret the R-sqaured value.

```{r}
ggplot(data = alc_hfi_2010, aes(x = hf_score, y = total_litres_of_pure_alcohol)) + geom_point()
```

## Is the amount of a country's alcohol consumption affected by how happy it's citizens are?

In my dataset is a happiness factor ranking of countries around the world. This happiness ranking is based on certain 
variables such as freedom of expression, freedom of religion and economic freedom.

## The happiness factor alcohol linear model
```{r mod_hf_alc}
(mod_hf_alc <- lm(total_litres_of_pure_alcohol ~ hf_score, data = alc_hfi_2010))
```

```{r summary-mod_hf_alc}
summary(mod_hf_alc)
```

```{r}
ggplot(data = alc_hfi_2010, aes(x = hf_score, y = total_litres_of_pure_alcohol)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```

You would think that countries with higher religious restrictions would translate to low amounts of alcohol consumed per person per year.  However, the plot above doesn't seem to support this assumption.  As you can see, there is a slight positive trend but the data really shows that there really is not.  Both positive and negative residuals are somewhat evenly distributed above and below the least of squares line and across the x-axis and are not tightly packed around the line.  This suggests a low correlation score which suggests a weak correlation to a positive linear trend. This assumption is supported by the very low R-squared value of less than 1% (0.009635).

## The personal freedom factor alcohol linear model
```{r mod_pf_alc}
(mod_pf_alc <- lm(total_litres_of_pure_alcohol ~ pf_score, data = alc_hfi_2010))
```

```{r summary-mod_pf_alc}
summary(mod_pf_alc)
```

```{r}
ggplot(data = alc_hfi_2010, aes(y = total_litres_of_pure_alcohol, x = pf_score)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```

## The economic factor alcohol linear model

```{r mod_ef_alc}
(mod_ef_alc <- lm(total_litres_of_pure_alcohol ~ ef_score, data = alc_hfi_2010))
```

```{r summary-mod_ef_alc}
summary(mod_ef_alc)
```

```{r}
ggplot(data = alc_hfi_2010, aes(y = total_litres_of_pure_alcohol, x = ef_score)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```


## Part 4 - Conclusion

### Why is the analysis important?
### Limitations of the analysis?

### References

### Appendix
#bar graph of top 10 soft skills as percentage
```{r plot bar graph of top 15 beer servings, echo=TRUE, fig.keep='all', message=FALSE, results='hide'}
head(alc_hfi_2010[order(-alc_hfi_2010$beer_servings),],15) %>% 
    ggplot(aes(y=reorder(countries,beer_servings),x=beer_servings,fill=countries)) +
      geom_bar(stat = 'identity',position=position_dodge()) +
      geom_text(aes(label=beer_servings), vjust=1.0, color="black", 
            position = position_dodge(0.9), size=3.0) +
      labs(x = ("Beer Servings"),y = ("Country"),
      title = ("Top 15 Countries 12 oz Beer Servings in 2010 per Person")  )
      theme_minimal()
```

#bar graph of top 10 soft skills as percentage
```{r plot bar graph of top 15 spirit servings,echo=TRUE, fig.keep='all', message=FALSE, results='hide'}
head(alc_hfi_2010[order(-alc_hfi_2010$spirit_servings),],15) %>% 
    ggplot(aes(y=reorder(countries,spirit_servings),x=spirit_servings,fill=countries)) +
      geom_bar(stat = 'identity',position=position_dodge()) +
      geom_text(aes(label=spirit_servings), vjust=1.0, color="black", 
            position = position_dodge(0.9), size=3.0) +
      labs(x = ("Spirit Servings"),y = ("Country"),
      title = ("Top 14 Countries Spirit Servings (1.5 oz) in 2010 per Person")  )
      theme_minimal()
```

#bar graph of top 10 soft skills as percentage
```{r plot bar graph of top 15 wine servings,echo=TRUE, fig.keep='all', message=FALSE, results='hide'}
head(alc_hfi_2010[order(-alc_hfi_2010$wine_servings),],15) %>% 
    ggplot(aes(y=reorder(countries,wine_servings),x=wine_servings,fill=countries)) +
      geom_bar(stat = 'identity',position=position_dodge()) +
      geom_text(aes(label=wine_servings), vjust=1.0, color="black", 
            position = position_dodge(0.9), size=3.0) +
      labs(x = ("wine Servings"),y = ("Country"),
      title = ("Top 14 Countries Wine (12 oz) Servings in 2010 per Person")  )
      theme_minimal()
```